---
title: "Table of regulated phosphosites for paper SILAC010-8"
author: "Marie Locard-Paulet"
date: "1 October 2014"
output: html_document
---

This document presents how the tables with regulated values for targets of interest were obtained.  
The tables used for this analysis are:  
- "SILAC1018TotV5_FAAposition.csv": all raw data from PD identification, with a column corresponding to the location of the first amino-acid of the identified sequence in the protein and a unique ID (`FirstAminoAcidPosition` and `UniqueID`, respectively). More interestingly, a normalisation was performed with the median of each sample before titanium enrichment, the resulting ratio Heavy/medium (log2 transformed) is presented in the column `normalisedratioHM` (with the columns `normalisedratioHM` and `NormalisationFactor`).  
- "Samples.csv" contains all information regarding samples, cell lines, fractions.  
In the "Tables folder":
- "HUVsignV5spF.csv", "MDAsignV5spF.csv" and "MDAsignV5spFLMNA.csv" contain the spectra corresponding to the list of regulated targets after manual inspection.  
- "HUVChecked.csv" and "MDAChecked.csv" are the equivalent tables after manual inspection of a broader range of phosphosites of interest.

Output tables with the median of log2(Heavy/Medium) value per phosphosite in HUVECs and MDA are stored in the folder "FinalTables".



```{r PackagesFunctions, cache=TRUE, echo=FALSE, message=FALSE, warning=FALSE, cache.path="cache_TableRegPsites/"}
# Packages:
library("ggplot2")
library("dplyr")

# Function:
FilledTable <- function(List, data.frame){
        mat <- matrix("", nrow=length(List), ncol=max(sapply(List, length)))
        for(x in 1:length(List))
                if(length(List[[x]]))
                {mat[x,1:length(List[[x]])] <- List[[x]]}
        cbind(data.frame, mat)
}
replace.vector <-function(x, tochange=unique(x), toreplace=1:length(unique(x))){
        for (y in 1:length(tochange))  x=replace(x, list=(x==tochange[y]), toreplace[y])
        x
}
tableCat <- function(inFrame) {
    outText <- paste(names(inFrame), collapse = " | ")
    outText <- c(outText, paste(rep("---", ncol(inFrame)), collapse = " | "))
    invisible(apply(inFrame, 1, function(inRow) {
        outText <<- c(outText, paste(inRow, collapse = " | "))
    }))
    return(outText)
}
```

First step: for each PSM, information regarding sample, cell line and fraction are retrieved from the "Samples.csv" file.


```{r InitialTable, cache=TRUE, echo=TRUE, warning=FALSE, cache.lazy=FALSE, cache.path="cache_TableRegPsites/"}
sample <- read.csv("Samples.csv")
tab <- read.csv("SILAC1018TotV5_FAAposition.csv")
tab <- tab[,c(1:ncol(tab)-1)] # remove the last column that contains only "NA".

m <- match(tab$SpectrumFile,sample$SpectrumFile)
df1 <- data.frame(tab, "Experiment"=sample$Experiment[m], "SampleName"=sample$SampleName[m], "Fraction"=sample$Fraction[m], "Cell"=sample$LabelledCell[m])
tab <- df1
tab <- tab[tab$qValue<=.01,]
tab <- tab[!is.na(tab$pRSSiteProbabilities),]
tab <- tab[tab$pRSSiteProbabilities!="",]
val1 <- length(unique(tab$Sequence))/length(unique(df1$Sequence))*100
Ambiguous <- tab[grepl(";", tab$ProteinGroupAccessions),]
UAmbSeq <- length(unique(Ambiguous$Sequence))
# write.csv(Ambiguous, "Ambiguous.csv", row.names=F) # if you want to have a 
# check at the ambiguous peptides (not attributable to a unique protein)
tab$ProteinGroupAccessions <- as.character(tab$ProteinGroupAccessions)
tab$ProteinGroupAccessions[(grepl("CLIP1_HUMAN", as.character(tab$ProteinDescriptions)))&(grepl("CLIP2_HUMAN", as.character(tab$ProteinDescriptions)))] <- "P30622"
tab$ProteinGroupAccessions[(grepl("TMCC2_HUMAN", as.character(tab$ProteinDescriptions)))&(grepl("TMCC3_HUMAN", as.character(tab$ProteinDescriptions)))] <- "O75069"
tab$FirstAminoAcidPosition <- as.character(tab$FirstAminoAcidPosition)
tab$FirstAminoAcidPosition[(grepl("CLIP1_HUMAN", as.character(tab$ProteinDescriptions)))&(grepl("CLIP2_HUMAN", as.character(tab$ProteinDescriptions)))] <- "346"
tab$FirstAminoAcidPosition[(grepl("TMCC2_HUMAN", as.character(tab$ProteinDescriptions)))&(grepl("TMCC3_HUMAN", as.character(tab$ProteinDescriptions)))] <- "436"
# Manual assessment of the "Ambiguous" table draw our attention of these two phosphopeptides. There log2(H/M) values may indicate phosphoregulation. Thus, I label them differently before I remove the ambiguous peptides from the analysis. These ones will be kept. This way, the localisation of the phosphosite will be based only on CLIP1 and TMCC2 sequences, respectively.
tab <- tab[!grepl(";", tab$ProteinGroupAccessions),]

##------
# Phosphosites ID:

prob <- as.character(tab$pRSSiteProbabilities)
P <- NULL
fp <- as.character(tab$FirstAminoAcidPosition)
for (i in 1:length(prob)){
        x <- strsplit(prob[i], split='; ', fix=T)[[1]]
        mat <- matrix(0, ncol=3,nrow=length(x))
        for (j in 1:length(x)){
                y <- strsplit(x[j], split=':', fix=T)[[1]]
                mat[j,3] <- y[2]
                y <- strsplit(y[1], split='(', fix=T)[[1]]
                y[2] <- strsplit(y[2], split=')', fix=T)[[1]][1]
                mat[j, 1:2] <- y
                mat[j,2] <- as.numeric(mat[j,2])+as.numeric(fp[i])-1
        }
        mat <- mat[as.numeric(mat[,3])>=60,,drop=F]
        mat <- mat[sort.list(as.numeric(mat[,3]),decreasing=T), ]
        mat <- as.vector(t(mat))
        P <- c(P, list(mat))
}
Uniprot <- as.character(tab$ProteinDescriptions)
Uniprot <- sapply(Uniprot, function(x) strsplit(x, split='[', fixed=T)[[1]][2])
Uniprot <- sapply(Uniprot, function(x) strsplit(x, split='_', fixed=T)[[1]][1])
names(Uniprot) <- NULL
Uniprot[grepl("CLIP2", tab$ProteinDescriptions)] <- "CLIP1/2"
Uniprot[(grepl("TMCC2", tab$ProteinDescriptions))&(grepl("TMCC3", tab$ProteinDescriptions))] <- "TMCC2/3"
# This way, the ambiguous proteins are considered as one. 
tab <- FilledTable(P, data.frame(tab, nSite=sapply(P,function(x) length(x)/3), Uniprot=Uniprot))
tab <- tab[(tab$nSite!=0), ] # Removes the peptides with ambiguous phosphosite position (pRS probability <= 60%).
ID <- vector(length=nrow(tab))
tab$Uniprot <- as.character(tab$Uniprot)
tab$"1" <- as.character(tab$"1")
tab$"2" <- as.numeric(as.character(tab$"2"))
tab$"4" <- as.character(tab$"4")
tab$"5" <- as.numeric(as.character(tab$"5"))
tab$"7" <- as.character(tab$"7")
tab$"8" <- as.numeric(as.character(tab$"8"))
for (i in 1:nrow(tab)){
        if (tab$nSite[i]%in%2) {ID[i] <- c(paste0(tab$Uniprot[i],'-',tab$"1"[i],tab$"2"[i],'-',tab$"4"[i],tab$"5"[i]))}
        if (tab$nSite[i]%in%3) {ID[i] <- c(paste0(tab$Uniprot[i],'-',tab$"1"[i],tab$"2"[i],'-',tab$"4"[i],tab$"5"[i],'-',tab$"7"[i],tab$"8"[i]))}
        else {ID[i] <- c(paste0(tab$Uniprot[i],'-',tab$"1"[i],tab$"2"[i]))}
}
tab <- data.frame(tab, ID=ID, stringsAsFactors = FALSE)

# Split HUVEC and MDA:

tab <- tab[!is.na(tab$log2HM),]
tMDA <- tab[tab$Cell=="MDA-MB-231",]
tHUV <- tab[tab$Cell=="HUVEC",] 
```


# HUVEC

```{r HUVEC, echo=TRUE, cache=TRUE, dependson="InitialTable", cache.lazy=FALSE, cache.path="cache_TableRegPsites/"}
tab <- tHUV

tab <- tab[!is.na(tab$log2HM),]
IDQuan <- paste0(tab$QuanResultID,"-",tab$SpectrumFile) # creates a unique quantification ID per phosphosite. 
tab <- data.frame(tab,IDQuan=IDQuan,stringsAsFactors=F)

QID <- lapply(unique(tab$IDQuan), function(y) tab[tab$IDQuan==y, 'X3']) 
names(QID) <- unique(tab$IDQuan) # list with the pRS probability of the first phosphosite for each QuanResultID
SID <- lapply(unique(tab$IDQuan), function(y) tab[tab$IDQuan==y, 'SpectrumFile']) 
names(SID) <- unique(tab$IDQuan) # list with the sample files for each QuanResultID
UID <- lapply(unique(tab$IDQuan), function(y) tab[tab$IDQuan==y, 'UniqueID']) 
names(UID) <- unique(tab$IDQuan) # list with the unique row ID for each QuanResultID


MaxpRS <- lapply(1:length(QID), function(x) aggregate(as.numeric(as.character(QID[[x]])), list(as.character(SID[[x]])), max)[,2]) 
names(MaxpRS) <- unique(tab$IDQuan) # determines which quantification result corresponds to the highest pRS probability for a phosphosite in a given sample.

v <- c(rep(0,length(MaxpRS)))
for (i in 1:length(MaxpRS)) {
        for (j in 1:length(UID[[i]])) {
                if (MaxpRS[[i]]==as.numeric(as.character(QID[[i]][j]))) {
                        v[i] <- as.character(UID[[i]][j])
                }
        }
}

tab1 <- tab[tab$UniqueID%in%v,] ## This table contain one quan per peptide + the peptide of this quanvalue with the max pRS

tab <- tab1
spchecked <- read.csv("Tables/HUVChecked.csv", as.is = T)
spchecked <- unique(spchecked$UniqueID[spchecked$X == "g"])
spchecked1 <- read.csv("Tables/HUVsignV5spF.csv", as.is = T)
spchecked1 <- unique(spchecked1$UniqueID[spchecked1$X == "g"])
spchecked <- c(spchecked, spchecked1)
tab <- tab[tab$UniqueID %in% spchecked,]

tab1 <- tab
#### Then, I perform quantification (average of all median per sample) on the phosphosites 

tab1$normalisedratioHM <- as.numeric(as.character(tab1$normalisedratioHM))
tab1$Experiment <- as.character(tab1$Experiment)
dat <- lapply(unique(tab1$ID), function(y) tab1[tab1$ID==y, 'normalisedratioHM'])
dat2 <- lapply(unique(tab1$ID), function(y) tab1[tab1$ID==y, 'Experiment'])
names(dat) <- unique(tab1$ID)

dat <- lapply(1:length(dat), function(x) aggregate(dat[[x]], list(as.character(dat2[[x]])), median)[,2]) # Keep the median of log2(H/M) per biological replicate in order to perform the t-test.
names(dat) <- unique(tab1$ID)

fc <- sapply(1:length(dat), function(y) mean(dat[[y]]))
Uniprot <- tab1$Uniprot[match(names(dat), tab1$ID)]

# Create matrix with median value per experiment for each phosphosite in the list:
mat <- matrix(ncol = (length(unique(tab1$Experiment)) + 1), nrow = length(dat))
mat[,1] <- names(dat)
colnames(mat) <- c("ID", unique(as.character(tab1$Experiment)))
for (i in 1:nrow(mat)) {
        dat2[[i]] <- unique(dat2[[i]])
        for (j in 1:length(dat[[i]])) {
                for (k in 2:(length(unique(tab1$Experiment)) + 1)) {
                        if (colnames(mat)[k] == dat2[[i]][j]) mat[i,k] <- dat[[i]][j]
                }
        }
}

tabHUV <- data.frame(mat, "FC" = fc, "Uniprot" = Uniprot)
write.csv(tabHUV, "FinalTables/TopHitsHUVECAverages.csv", row.names = F)
```

The table of regulated phosphosites in the HUVEC is:

```{r tableHUVEC, cache=TRUE, echo=FALSE, results='asis', dependson="HUVEC", cache.path="cache_TableRegPsites/"}
tabHUV2 <- tableCat(tabHUV[order(tabHUV$FC),])
cat(tabHUV2, sep = "\n")
```



```{r HUVECfigure, cache=TRUE, echo=FALSE, fig.align='center', cache=TRUE, fig.height=7, fig.width=9, dependson="HUVEC", cache.path="cache_TableRegPsites/"}
# Now, I want to have a figure with the values: I plot "normalisedlog2HM" from tab1 (filtered table with only checked spectra):
v <- vector(length = nrow(tab1))
v <- tabHUV$FC[match(tab1$ID, tabHUV$ID)]
tab1 <- data.frame(tab1, "FC" = v, stringsAsFactors = F)

tab1 <- tab1[order(tab1$FC),]
tab1$ID <- factor(as.character(tab1$ID), ordered(as.character(tab1$ID[!duplicated(tab1$ID)])))

g <- ggplot(data = tab1, aes(x = ID, y = normalisedratioHM, shape = Experiment, col = Fraction)) + geom_hline(yintercept=0,) + geom_hline(yintercept=-1,linetype="dashed",col="grey") + geom_hline(yintercept=1,linetype="dashed",col="grey")  
g + geom_point() + theme_bw()  + theme(axis.text.x=element_text(angle=-90), axis.title.x=element_blank()) + labs(title="Checked phosphosite in the HUVEC", y="Fold Change (Log2)")  + scale_colour_manual(values = c("membrane"="springgreen4", "cytoplasm"="darkorange2"))
```
Figure with only the phosphosites with an average fold change of more than 1.5:

```{r HUVECfigureFC, cache=TRUE, echo=FALSE, fig.align='center', cache.path="cache_TableRegPsites/", fig.height=7, fig.width=9, dependson="HUVEC"}

tab1 <- tab1[tab1$ID %in% tabHUV$ID[(tabHUV$FC>=0.5) | (tabHUV$FC<=-0.5)],]
v <- vector(length = nrow(tab1))
v <- tabHUV$FC[match(tab1$ID, tabHUV$ID)]
tab1 <- data.frame(tab1, "FC" = v, stringsAsFactors = F)

tab1 <- tab1[order(tab1$FC),]
tab1$ID <- factor(as.character(tab1$ID), ordered(as.character(tab1$ID[!duplicated(tab1$ID)])))

g <- ggplot(data = tab1, aes(x = ID, y = normalisedratioHM, shape = Experiment, col = Fraction)) + geom_hline(yintercept=0,) + geom_hline(yintercept=-1,linetype="dashed",col="grey") + geom_hline(yintercept=1,linetype="dashed",col="grey")  
g + geom_point() + theme_bw()  + theme(axis.text.x=element_text(angle=-90), axis.title.x=element_blank()) + labs(title="Regulated phosphosite in the HUVEC", y="Fold Change (Log2)")  + scale_colour_manual(values = c("membrane"="springgreen4", "cytoplasm"="darkorange2"))
```
# MDA

```{r MDA, echo=TRUE, cache=TRUE, dependson="InitialTable", cache.lazy=FALSE, cache.path="cache_TableRegPsites/"}
tab <- tMDA
tab <- tab[!is.na(tab$log2HM),]
IDQuan <- paste0(tab$QuanResultID,"-",tab$SpectrumFile) # creates a unique quantification ID per phosphosite. 
tab <- data.frame(tab,IDQuan=IDQuan,stringsAsFactors=F)

QID <- lapply(unique(tab$IDQuan), function(y) tab[tab$IDQuan==y, 'X3']) 
names(QID) <- unique(tab$IDQuan) # list with the pRS probability of the first phosphosite for each QuanResultID
SID <- lapply(unique(tab$IDQuan), function(y) tab[tab$IDQuan==y, 'SpectrumFile']) 
names(SID) <- unique(tab$IDQuan) # list with the sample files for each QuanResultID
UID <- lapply(unique(tab$IDQuan), function(y) tab[tab$IDQuan==y, 'UniqueID']) 
names(UID) <- unique(tab$IDQuan) # list with the unique row ID for each QuanResultID


MaxpRS <- lapply(1:length(QID), function(x) aggregate(as.numeric(as.character(QID[[x]])), list(as.character(SID[[x]])), max)[,2]) 
names(MaxpRS) <- unique(tab$IDQuan) # determines which quantification result corresponds to the highest pRS probability for a phosphosite in a given sample.

v <- c(rep(0,length(MaxpRS)))
for (i in 1:length(MaxpRS)) {
        for (j in 1:length(UID[[i]])) {
                if (MaxpRS[[i]]==as.numeric(as.character(QID[[i]][j]))) {
                        v[i] <- as.character(UID[[i]][j])
                }
        }
}


tab <- tab[tab$UniqueID%in%v,] ## This table contain one quan per peptide + the peptide of this quanvalue with the max pRS


spchecked <- read.csv("Tables/MDAChecked2.csv", as.is = T)
spchecked <- unique(spchecked$UniqueID[spchecked$X == "g"])
spchecked1 <- read.csv("Tables/MDAsignV5spF.csv", as.is = T)
spchecked1 <- unique(spchecked1$UniqueID[spchecked1$X == "g"])
spchecked3 <- read.csv("Tables/MDAsignV5spFLMNA.csv", as.is = T)
spchecked3 <- unique(spchecked3$UniqueID[spchecked3$X == "g"])
spchecked <- c(spchecked, spchecked1, spchecked3)
tab1 <- tab[tab$UniqueID %in% spchecked,]



#### Then, I perform quantification (average of all median per sample) on the phosphosites 

tab1$normalisedratioHM <- as.numeric(as.character(tab1$normalisedratioHM))
tab1$Experiment <- as.character(tab1$Experiment)
dat <- lapply(unique(tab1$ID), function(y) tab1[tab1$ID==y, 'normalisedratioHM'])
dat2 <- lapply(unique(tab1$ID), function(y) tab1[tab1$ID==y, 'Experiment'])
names(dat) <- unique(tab1$ID)

dat <- lapply(1:length(dat), function(x) aggregate(dat[[x]], list(as.character(dat2[[x]])), median)[,2]) # Keep the median of log2(H/M) per biological replicate in order to perform the t-test.
names(dat) <- unique(tab1$ID)

fc <- sapply(1:length(dat), function(y) mean(dat[[y]]))
Uniprot <- tab1$Uniprot[match(names(dat), tab1$ID)]

# Create matrix with median value per experiment for each phosphosite in the list:
mat <- matrix(ncol = (length(unique(tab1$Experiment)) + 1), nrow = length(dat))
mat[,1] <- names(dat)
colnames(mat) <- c("ID", unique(as.character(tab1$Experiment)))
for (i in 1:nrow(mat)) {
        dat2[[i]] <- unique(dat2[[i]])
        for (j in 1:length(dat[[i]])) {
                for (k in 2:(length(unique(tab1$Experiment)) + 1)) {
                        if (colnames(mat)[k] == dat2[[i]][j]) mat[i,k] <- dat[[i]][j]
                }
        }
}

tabMDA <- data.frame(mat, "FC" = fc, "Uniprot" = Uniprot)
write.csv(tabMDA, "FinalTables/TopHitsMDAAverages.csv", row.names = F)
```

The table of regulated phosphosites in the MDA is:

```{r tableMDA, cache=TRUE, echo=FALSE, results='asis', dependson="MDA", cache.path="cache_TableRegPsites/"}
tabMDA2 <- tableCat(tabMDA[order(tabMDA$FC),])
cat(tabMDA2, sep = "\n")
```


```{r MDAfigure, cache=TRUE, echo=FALSE, fig.align='center', fig.height=7, fig.width=9, dependson="MDA", cache.path="cache_TableRegPsites/"}
# Now, I want to have a figure with the values: I plot "normalisedlog2HM" from tab1 (filtered table with only checked spectra):
v <- vector(length = nrow(tab1))
v <- tabMDA$FC[match(tab1$ID, tabMDA$ID)]
tab1 <- data.frame(tab1, "FC" = v, stringsAsFactors = F)

tab1 <- tab1[order(tab1$FC),]
tab1$ID <- factor(as.character(tab1$ID), ordered(as.character(tab1$ID[!duplicated(tab1$ID)])))

g <- ggplot(data = tab1, aes(x = ID, y = normalisedratioHM, shape = Experiment, col = Fraction)) + geom_hline(yintercept=0,) + geom_hline(yintercept=-1,linetype="dashed",col="grey") + geom_hline(yintercept=1,linetype="dashed",col="grey")  
g + geom_point() + theme_bw()  + theme(text = element_text(size=7), axis.text.x=element_text(angle=-90), axis.title.x=element_blank()) + labs(title="Checked phosphosite in the MDA", y="Fold Change (Log2)")  + scale_colour_manual(values = c("membrane"="springgreen4", "cytoplasm"="darkorange2"))
```
Only the ones with a fold change of at least 1.5:
```{r MDAfigureFC, cache=TRUE, echo=FALSE, fig.align='center', cache.path="cache_TableRegPsites/", fig.height=7, fig.width=9, dependson="MDA"}

tab1 <- tab1[tab1$ID %in% tabMDA$ID[(tabMDA$FC>=0.5) | (tabMDA$FC<=-0.5)],]
v <- vector(length = nrow(tab1))
v <- tabMDA$FC[match(tab1$ID, tabMDA$ID)]
tab1 <- data.frame(tab1, "FC" = v, stringsAsFactors = F)

tab1 <- tab1[order(tab1$FC),]
tab1$ID <- factor(as.character(tab1$ID), ordered(as.character(tab1$ID[!duplicated(tab1$ID)])))

g <- ggplot(data = tab1, aes(x = ID, y = normalisedratioHM, shape = Experiment, col = Fraction)) + geom_hline(yintercept=0,) + geom_hline(yintercept=-1,linetype="dashed",col="grey") + geom_hline(yintercept=1,linetype="dashed",col="grey")  
g + geom_point() + theme_bw()  + theme(text = element_text(size=8), axis.text.x=element_text(angle=-90), axis.title.x=element_blank()) + labs(title="Regulated phosphosite in the MDA", y="Fold Change (Log2)")  + scale_colour_manual(values = c("membrane"="springgreen4", "cytoplasm"="darkorange2"))
```

